---
title: "Model Factor Scores"
author: "Robin"
date: "Feb 6 2020"
output: 
  html_document:
    toc: yes
    fig_width: 6
    fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo = FALSE, cache = FALSE, include = FALSE}
# load libraries
library(lme4)
library(lmerTest)
library(AICcmodavg)
library(bbmle)
library (jtools)
library(ggstance)
library(cowplot)
library(sjstats)
library(dplyr)
library(readr)
library(broom)
library(insight)
library(glmmTMB)
library(ggplot2)
```

```{r, cache = TRUE}
# Read data 
setwd('invariance-repetitive-bx/') # change to the correct wdir
dat = read.csv('data/mnlfa_output.csv')
```


```{r}
# Cohort and male/female breakdown
dat %>%
  group_by(TABLET, MALE) %>%
  summarize(n=n_distinct(ID2))
```


# Plot distributions
## Composite RBS-EC scores
Distribution of Composite RBS-EC scores (n=606). All 34 items included. Mean = `r round(mean(dat$comp_items, na.rm = TRUE),2)`, SD = `r round(sd(dat$comp_items, na.rm=TRUE),2)`.
```{r cache = TRUE}
p1 = ggplot(data = dat, aes(x = comp_items)) + 
  geom_histogram(color = 'black', binwidth = 1) + 
  scale_x_continuous(breaks = seq(0,34,2)) +
  xlab('Composite RBS-EC Score') +
  ylab('Frequency') +
  theme(text = element_text(size=16)) +
  theme_bw()
```

## Subscales
```{r}
#  Rep Motor 
p1 = ggplot(data = dat, aes(x = as.numeric(REP_ETA1))) + 
  geom_histogram(color = 'black', binwidth = 0.5,
                 aes(y = ..density..) ) +
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(dat$REP_ETA2), 
                          sd=sd(dat$REP_ETA2))) +
  xlab('Repetitive Motor (Factor Scores)') +
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3,3,0.5)) +
  theme_bw() + 
  theme(text = element_text(size=16))
  
p2 = ggplot(data = dat, aes(x = as.numeric(repmean))) + 
  geom_histogram(color = 'black', binwidth = .1,
                 aes(y = ..density..) ) +
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(dat$repmean), 
                          sd=sd(dat$repmean))) +
  xlab('Repetitive Motor (Raw Mean Scores)') +
  scale_x_continuous(limits = c(-.1, 1.1), breaks = seq(-0,1,.1)) +
  theme_bw() +
  theme(text = element_text(size=16))

p3 = gridExtra::grid.arrange(p1, p2, nrow = 2)
p3
# SELF 
p1 = ggplot(data = dat, aes(x = SELF_ETA1)) + 
    geom_histogram(color = 'black', binwidth =0.5,
                   aes(y = ..density..) ) +
    stat_function(fun=dnorm,
                  color="red",
                  args=list(mean=mean(dat$SELF_ETA1), 
                            sd=sd(dat$SELF_ETA1))) +
  xlab('Self-directed (Factor Scores)') +
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3,3,0.5)) +
  theme_bw() +theme(text = element_text(size=16))
p2 = ggplot(data = dat, aes(x =selfmean)) + 
  geom_histogram(color = 'black',  binwidth = .2,
                 aes(y = ..density..) ) +
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(dat$selfmean), 
                          sd=sd(dat$selfmean))) +
  xlab('Self-directed (Raw Mean Scores)') +
    scale_x_continuous(limits = c(-0.1, 1.1), breaks = seq(-0,1,0.2)) + 
  theme_bw() +theme(text = element_text(size=16))
p3 = gridExtra::grid.arrange(p1, p2, nrow = 2)
p3

p1 = ggplot(data = dat, aes(x = as.numeric(RIT_RES_ETA1))) + 
  geom_histogram(color = 'black', binwidth = 0.5,
                 aes(y = ..density..) ) +
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(dat$RIT_RES_ETA1), 
                          sd=sd(dat$RIT_RES_ETA1))) +
  xlab('Higher-order (Factor Scores)') + 
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3,3,0.5))+
  theme_bw() +
  theme(text = element_text(size=16))

p2= ggplot(data = dat, aes(x = as.numeric(resritmean))) + 
  geom_histogram(color = 'black', binwidth = 0.1,
                 aes(y = ..density..) ) +
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(dat$resritmean), 
                          sd=sd(dat$resritmean))) +
  xlab('Higher-order (Raw mean scores)') + 
  scale_x_continuous(limits = c(-.1, 1.1), breaks = seq(0,1,0.1))+
  theme_bw() +
  theme(text = element_text(size=16))
p3 = gridExtra::grid.arrange(p1, p2, nrow = 2)
p3
```


# Growth curve analyses
I fit models sequentially in the following steps:  
1. Unconditional (intercept-only) model.  
2. Determine the functional form - linear or quadratic? Fixed or random effects?  
3. Test for effects of Sex/Cohort on Intercept and Slope  
4. Remove n.s. covariates. Select final model.

## Repetitive Motor 
### Repetitive Motor - Factor scores
1. <b>Functional form - fixed effects. </b>.
```{r, echo = TRUE}

lmer.0=lmer(REP_ETA1~ 1  +( 1|ID2), data = dat, REML = FALSE)
lmer.1=lmer(REP_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
lmer.2=lmer(REP_ETA1~ 1  + AGE18+ I(AGE18^2) + ( 1 |ID2), data = dat, REML = FALSE)
```
```{r}
# Model comparison
mynames <- as.character (c("lmer.0", "lmer.1", "lmer.2")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.0),
                 logLik(lmer.1),
                 logLik(lmer.2),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc #dAICc over 2 means significantly better
```

We select lmer.1. While a quadratic FE of age has a lower AICc, dAICc of 0.2 does not justify the increase in model complexity.  
2. <b>Functional form - random effects</b>. Adding a linear RE of age improves model fit.
```{r echo = TRUE}
lmer.3 = lmer(REP_ETA1~ 1  + AGE18+  ( 1 +AGE18 |ID2), data = dat, REML = FALSE)
```
```{r cache = TRUE}
mynames <- as.character (c("lmer.1", "lmer.3")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.1),
                 logLik(lmer.3),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc$rweight <- max(myaicc$weight) / myaicc$weight
myaicc #dAI


```
We select model with fixed and random linear effect of age (lmer.3).  Change in log-likelihood = `r logLik(lmer.3)[[1]] - logLik(lmer.1)[[1]] `.   

3. <b>Selecting covariates</b>.
```{r, echo = TRUE}
# 3. Assess covariates (& their interaction with age)
lmer.4 =  lmer(REP_ETA1~ 1  + AGE18+  MALE + TABLET +
                 AGE18*MALE + AGE18*TABLET + ( 1 +AGE18 |ID2), data = dat, REML = FALSE)
summary(lmer.4)
# drop n.s. covariates
lmer.4b =  lmer(REP_ETA1~ 1  + AGE18+  MALE + 
                 AGE18*MALE +  ( 1 +AGE18 |ID2), data = dat, REML = FALSE)
```

There is no significant effect on cohort (TABLET) on intercept or slope. Remove from model and confirm that this does not significantly impact model fit. 
```{r}
anova(lmer.4, lmer.4b) # confirm that lmer.4 does not have significantly lower AIC

mynames <- as.character (c("lmer.3", "lmer.4b")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.3),
                 logLik(lmer.4),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc$rweight <- max(myaicc$weight) / myaicc$weight
myaicc 
```

5. <b> Final Model </b>.  We select lmer.4b.
```{r echo = TRUE}
summary(lmer.4b)
```

```{r}
# Pseudo-r^2
dat$repyhat = predict(lmer.4b, newdata=dat, re.form = NA )
R = cor(x = dat$repyhat, 
        y =dat$REP_ETA1, use = "complete.obs")

# 5. Scaling on within-person variance
#summary(lmer.0)
l1_sd = sqrt(1.1064)#  l1-var from random intercept model 
age_eff = -0.109913 # summary(lmer.1)
mo_18 =  0.039220 # intercept
mo_8 =  -10*age_eff # For a ten unique DECREASE in age (to get to 8 mos)
mo_37 =  19*age_eff

```
Pseudo-r2 of final model= `r R^2`    
Scaling on within-person variance, age effect translates to `r (mo_37-mo_8) / l1_sd` change in behaviors from 8-37 months. 

### Rep - Raw mean scores 
We use a generalized mixed-effects linear model with a logistic linking function bc of the data skew. 
1. <b> Functional form - fixed effects </b>  
```{r, echo = TRUE}
glm.0 = glmer(repmean ~ 1 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.1 = glmer(repmean ~ 1 + AGE18 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.2 = glmer(repmean ~ 1 + AGE18 + I(AGE18^2) + (1|ID2), data = dat, family = binomial(link = "logit"))
```
```{r}
mynames <- as.character (c("glm.0", "glm.1", "glm.2")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.0),
                 logLik(glm.1),
                 logLik(glm.2),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc$rweight <- round(max(myaicc$weight) / myaicc$weight,2)
myaicc 
```
We select linear fixed effect. dAICc does not justify quadratic term.

<b> Functional form - random effects </b>. Adding in a linear RE of age improves model fit 
```{r echo = TRUE}
glm.3 = glmer(repmean ~ 1 + AGE18 +(1 + AGE18|ID2), data = dat, family = binomial(link = "logit"))
```

```{r}
mynames <- as.character (c("glm.1", "glm.3")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.1),
                 logLik(glm.3),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc$rweight <- round(max(myaicc$weight) / myaicc$weight,2)
myaicc

```
We select model with fixed and random linear effect of age (glm.3).  Change in log-likelihood = `r logLik(glm.3)[[1]] - logLik(glm.1)[[1]] `.  

3. <b> Selecting covariates </b>. Models fails to converge
```{r echo = TRUE, warning = TRUE}
glm.4 = glmer(repmean ~ 1 + AGE18 + 
                MALE + TABLET + AGE18:MALE + AGE18:TABLET +(1 + AGE18|ID2), data = dat, 
              family = binomial(link = "logit"))
glm.4b = glmer(repmean ~ 1 + AGE18 + 
                MALE + TABLET  +(1 + AGE18|ID2), data = dat, 
              family = binomial(link = "logit"))
```

4. <b> Final Model Selection </b>. glm.3. 
```{r echo = TRUE}
summary(glm.3)
```
```{r}
dat$rep_meanyhat = predict(glm.3, newdata=dat, re.form = NA )

R = cor(x = dat$rep_meanyhat, 
        y =dat$repmean, use = "complete.obs")
```
Pseudo-r2 of final model = `r R^2`.


### Repetitive Motor - Figures 
Raw mean trajectory - y-axis in probability.
```{r}
glm.3 = glmer(repmean ~ 1 + AGE18 +(1 + AGE18|ID2), data = dat, family = binomial(link = "logit"))
dat$rep_ymean = predict(glm.3, newdat = dat, type = 'response')   
# logit2prob <- function(logit){
#   odds <- exp(logit)
#   prob <- odds / (1 + odds)
#   return(prob)
# }
# PLOT THE RAW MEAN, with model estimated probability overlayed 
p1 = ggplot(data = dat, aes(x = AGE, y = repmean)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = rep_ymean), color = 'black') + 
  ylab('Repetitive Motor - Raw Mean Scores') +
  xlab('Age (months)')  +
  scale_y_continuous(limits = c(mean(dat$repmean) - 2.5*sd(dat$repmean),
                               mean(dat$repmean) + 2.5*sd(dat$repmean))) +
  theme_bw() + 
  theme(text = element_text(size=16)) +
  scale_x_continuous(breaks = seq(9, 36,3))
p1
```
MNLFA score
```{r cache=TRUE}
lmer.6 = lmer(REP_ETA1~ 1  + AGE18+  MALE +
                 AGE18*MALE +  ( 1 +AGE18 |ID2), data = dat, REML = FALSE)

dat$rep_yhat_eta = predict(lmer.6, newdat = dat)   

p2 = ggplot(data = dat, aes(x = AGE, y = REP_ETA1)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = rep_yhat_eta),color = 'black') + 
  ylab('Repetitive Motor - MNFLA Scores') +
  xlab('Age (months)') +
  scale_y_continuous(limits = c(mean(dat$REP_ETA1) - 2.5*sd(dat$REP_ETA1),
                                mean(dat$REP_ETA1) + 2.5*sd(dat$REP_ETA1))) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))

# ggsave(p1, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_rep_mean.jpg', height = 4, width=7)
# ggsave(p2, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_rep_eta.jpg', height = 4, width=7)
p2
```
Age x Sex interaction
```{r cache = TRUE}
dat$male2 = ifelse(dat$MALE == 1, "Male", "Female")
dat$male2 = as.factor(dat$male2)
p3= ggplot(data = dat, aes(x = AGE, y = REP_ETA1, group_by(male2)), color =male2) + 
  geom_smooth(aes(y = rep_yhat_eta, color = male2), method = 'lm') + 
  ylab('Repetitive Motor - MNFLA Scores') +
  xlab('Age (months)') +
  scale_y_continuous(limits = c(mean(dat$REP_ETA1) - 2.5*sd(dat$REP_ETA1),
                                mean(dat$REP_ETA1) + 2.5*sd(dat$REP_ETA1))) +
  scale_color_discrete(name = 'Sex', labels = c('Female', 'Male')) + 
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p3
# ggsave(p3, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_rep_eta_sexinteraction.jpg', height = 4, width=7)
```

## Self-directed
### Self-Directed - Factor scores 
1. <b> Functional form - Fixed effects </b>
```{r echo = TRUE}
lmer.0=lmer(SELF_ETA1~ 1  +( 1|ID2), data = dat, REML = FALSE)
lmer.1=lmer(SELF_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
lmer.2=lmer(SELF_ETA1~ 1  + AGE18+ I(AGE18^2) + ( 1 |ID2), data = dat, REML = FALSE)
```

lmer.1 (linear FE of age) selected as best-fitting model.
```{r}
mynames <- as.character (c("lmer.0", "lmer.1", "lmer.2")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.0),
                 logLik(lmer.1),
                 logLik(lmer.2),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc
```
2. <b> Functional form-random effects. </b>
```{r echo = TRUE}
lmer.1=lmer(SELF_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
lmer.1b=lmer(SELF_ETA1~ 1  + AGE18+ ( 1 + AGE18|ID2), data = dat, REML = FALSE)
```

```{r}
mynames <- as.character (c("lmer.1", "lmer.1b")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.1),
                 logLik(lmer.1b),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 
```
Adding RE of age does not improve model fit. We go with intercept-only for RE.

3. <b> Testing covariates </b>
```{r echo = TRUE}
lmer.2=lmer(SELF_ETA1~ 1  + AGE18+ MALE + TABLET  + 
              AGE18*MALE + AGE18*TABLET + ( 1|ID2), data = dat, REML = FALSE)
summary(lmer.2)
```
No significant covariates. Confirm that removing does not hurt model fit.
```{r}
anova(lmer.1, lmer.2)
```

4. <b> Final Model </b>
```{r echo = TRUE}
lmer.1=lmer(SELF_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
```


```{r}
dat$self_yhat = predict(lmer.1, newdata=dat, re.form = NA )
R = cor(x = dat$self_yhat, 
        y =dat$SELF_ETA1, use = "complete.obs")
#scale on w/in person variance 
l1_sd = sqrt(0.5903)#  l1-var from random intercept model (lmer.0)
age_eff = -0.056794 # lmer.1
mo_8 =  -10*age_eff # For a ten unique DECREASE in age (to get to 8 mos)
mo_37 =  19*age_eff



```
Pseudo-r2 of final model = `r R^2`.  
Scaling on within-person l1 variability, the age affect translates to a `r (mo_37-mo_8) / l1_sd` change in behaviors from 8-37 months. 

### Self-Dir - Raw Mean
1. <b> Functional form - fixed effects </b>
```{r echo = TRUE, warnings=TRUE}
# 1. Functional form - fixed effects 
glm.0 = glmer(selfmean ~ 1 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.1 = glmer(selfmean ~ 1 + AGE18 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.2 = glmer(selfmean ~ 1 + AGE18 + I(AGE18^2) + (1|ID2), data = dat, family = binomial(link = "logit")) # did not converge 
```

```{r}
mynames <- as.character (c("glm.0", "glm.1")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.0),
                 logLik(glm.1),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 
```
Linear FE of age improves model fit from intercept-only model. 

2. <b> Functional form - random effects </b>
```{r echo = TRUE}
glm.1b=glmer(selfmean ~ 1 + AGE18 + (1+AGE18|ID2), data = dat, family = binomial(link = "logit"))
```

```{r}
mynames <- as.character (c("glm.1", "glm.1b")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.1),
                 logLik(glm.1b),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 
```
Adding random effect of age does not improve model fit. We include intercept-only as random effect.

3. <b> Adding covariates </b>
```{r echo = TRUE}
glm.2=glmer(selfmean ~ 1 + AGE18 + 
              MALE + TABLET + AGE18*TABLET + AGE18*MALE+(1|ID2), data = dat, family = binomial(link = "logit")) 
summary(glm.2)
```
Remove n.s. covariates & confirm it does not hurt model fit. 
```{r echo = TRUE}
glm.3=glmer(selfmean ~ 1 + AGE18 + 
              TABLET + (1|ID2), data = dat, family = binomial(link = "logit")) 
anova(glm.2, glm.3)
```

```{r}
mynames <- as.character (c("glm.1","glm.3")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.1),
                 logLik(glm.3),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 
```
Model w/ tablet covariate has  better fit.  

4. <b> Final model </b>
```{r echo = TRUE}
glm.3=glmer(selfmean ~ 1 + AGE18 + 
              TABLET + (1|ID2), data = dat, family = binomial(link = "logit")) 
```


### Self-Dir - Figures
Raw mean trajectory
```{r}
glm.3=glmer(selfmean ~ 1 + AGE18 + 
              TABLET + (1|ID2), data = dat, family = binomial(link = "logit")) 
dat$self_yhat_mean = predict(glm.3, newdat = dat, type = 'response')   
p2 = ggplot(data = dat, aes(x = AGE, y = selfmean)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = self_yhat_mean),color = 'black') + 
  ylab('Self-directed  Mean Scores') +
  xlab('Age (months)') +
  theme_bw() + 
  theme(text = element_text(size=16)) +
  scale_y_continuous(limits = c(mean(dat$selfmean) - 3*sd(dat$selfmean),
                                mean(dat$selfmean) + 3*sd(dat$selfmean))) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p2
```
Factor score trajectory
```{r}
lmer.1=lmer(SELF_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
dat$self_yhat_eta = predict(lmer.1, newdat = dat)

p1 = ggplot(data = dat, aes(x = AGE, y = SELF_ETA1)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = self_yhat_eta), color = 'black') + 
  ylab('Self-directed Factor Scores') +
  xlab('Age (months)') +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_y_continuous(limits = c(mean(dat$SELF_ETA1) -3*sd(dat$SELF_ETA1),
                                mean(dat$SELF_ETA1) + 3*sd(dat$SELF_ETA1))) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p1
# ggsave(p1, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_self_eta.jpg', height = 4, width=7)
# ggsave(p2, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_self_mean.jpg', height = 4, width=7)
```

## Higher-order
### Higher - Factor scores 
1. <b> Functional form - fixed effects </b>
```{r echo = TRUE} 
lmer.0=lmer(RIT_RES_ETA1~ 1  +( 1|ID2), data = dat, REML = FALSE)
lmer.1=lmer(RIT_RES_ETA1~ 1  + AGE18+ ( 1|ID2), data = dat, REML = FALSE)
lmer.2=lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + ( 1 |ID2), data = dat, REML = FALSE)
```
```{r}
mynames <- as.character (c("lmer.0", "lmer.1", "lmer.2")) 
myaicc  <- ICtab(logLik(lmer.0),
                 logLik(lmer.1),
                 logLik(lmer.2),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 
```
Quadratic FE of age significantly improves model fit.  
2. <b> Functional form - random effects </b>
```{r echo = TRUE}
lmer.3=lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + ( 1 + AGE18 |ID2), data = dat, REML = FALSE)
lmer.4=lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + ( 1  + AGE18 + I(AGE18^2)|ID2), data = dat, REML = FALSE) # failed to converge - not a candidate model
```
```{r}
mynames <- as.character (c("lmer.2", "lmer.3")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.2),
                 logLik(lmer.3),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc 

```
Adding RE of age signigicantly improves model fit. Change in log likelihood = `r logLik(lmer.1)[[1]] - logLik(lmer.3)[[1]]`.

3. <b> Test covariates </b>
```{r echo = TRUE}
# Test covariates, and interaction with linear & quadratic age 
lmer.5=lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + MALE + TABLET +
              AGE18*MALE + AGE18*TABLET + 
               I(AGE18^2)*MALE +  I(AGE18^2)*TABLET+
              ( 1 + AGE18 |ID2), data = dat, REML = FALSE)
summary(lmer.5)
```

Remove n.s. covariates & confirm it does not impact model fit
```{r}
lmer.5b = lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + MALE +
              AGE18*MALE  + ( 1 + AGE18 |ID2), data = dat, REML = FALSE)
anova(lmer.5, lmer.5b) # doesn't hurt model fit to drop n.s. covariates 
```

```{r}
mynames <- as.character (c("lmer.3", "lmer.5b")) # model numbers as characters
myaicc  <- ICtab(logLik(lmer.3),
                 logLik(lmer.5b),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc # We pick lmer.5b (includes interactoin with linear age and cohort)

```

4. <b> Select final model </b>
```{r echo = TRUE}
lmer.5b = lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + MALE +
              AGE18*MALE  + ( 1 + AGE18 |ID2), data = dat, REML = FALSE)


```

```{r}
# Pseudo-R^2
dat$resrit_yhat_eta = predict(lmer.5b, newdat = dat)
R = cor(x = dat$resrit_yhat_eta, 
        y =dat$RIT_RES_ETA1, use = "complete.obs")


l1_sd = sqrt(0.4107 )#  l1-var from random intercept model
age_eff = -2.134e-03 # quadratic age eff
mo_18 = -2.151e-03 # Intercept 
mo_37 =  19*age_eff

```
R2 of final model = `r R^2`.  Scaling on within-person l1 variance, this results in a `r (mo_37-mo_18) / l1_sd` SD change in behavior from 18-37 months.

### Higher - raw mean 
1. <b> Functional form - fixed effects </b>
```{r echo = TRUE}
glm.0 = glmer(resritmean ~ 1 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.1 = glmer(resritmean ~ 1 + AGE18 + (1|ID2), data = dat, family = binomial(link = "logit"))
glm.2 = glmer(resritmean ~ 1 + AGE18 + I(AGE18^2) + (1|ID2), data = dat, family = binomial(link = "logit"))
```
```{r}
mynames <- as.character (c("glm.0", "glm.1", "glm.2")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.0),
                 logLik(glm.1),
                 logLik(glm.2),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc
```
We adopt intercept-only model. 

2. <b> Test for covariates </b>
```{r}
glm.3 =  glmer(resritmean ~ 1 + MALE + TABLET+ (1|ID2), data = dat, family = binomial(link = "logit"))
```

```{r}
mynames <- as.character (c("glm.0", "glm.3")) # model numbers as characters
myaicc  <- ICtab(logLik(glm.0),
                 logLik(glm.3),
                 type ="AICc",
                 nobs = nrow(na.omit(dat)),
                 weights =TRUE,delta =TRUE,sort=TRUE,mnames=mynames,base=TRUE)
myaicc # Intercept-only model is better 
```
Adding covariates do not improve model fit. 

### Higher-order - Figures
Raw mean trajectory
```{r}
glm.0 = glmer(resritmean ~ 1 + (1|ID2), data = dat, family = binomial(link = "logit"))
dat$resrit_yhat_mean = predict(glm.0, newdat = dat, type = 'response')   
p2 = ggplot(data = dat, aes(x = AGE, y = resritmean)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = resrit_yhat_mean),color = 'black') + 
  ylab('Higher-order Mean Scores') +
  xlab('Age (months)') +
  theme_bw() + 
  theme(text = element_text(size=16)) +
  scale_y_continuous(limits = c(mean(dat$resritmean) - 4*sd(dat$resritmean),
                                mean(dat$resritmean) + 4*sd(dat$resritmean))) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p2

```
Factor-score trajectory
```{r}
lmer.5 = lmer(RIT_RES_ETA1~ 1  + AGE18+ I(AGE18^2) + MALE +
              AGE18*MALE + ( 1 + AGE18 |ID2), data = dat, REML = FALSE)

dat$resrit_yhat_eta = predict(lmer.3, newdat = dat)

p1 = ggplot(data = dat, aes(x = AGE, y = RIT_RES_ETA1)) + 
  geom_jitter(alpha = .4, color = 'indianred') + 
  geom_smooth(aes(y = resrit_yhat_eta), color = 'black') + 
  ylab('Higher-order Factor Scores') +
  xlab('Age (months)') +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_y_continuous(limits = c(mean(dat$RIT_RES_ETA1) -4*sd(dat$RIT_RES_ETA1),
                                mean(dat$RIT_RES_ETA1) + 4*sd(dat$RIT_RES_ETA1))) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p1



# ggsave(p1, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_resrit_eta.jpg', height = 4, width=7)
# ggsave(p2, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_resrit_mean.jpg', height = 4, width=7)
# ggsave(p3, filename = '/Users/sifre002/Box/sifre002/11_Figures/Longit_RBS_MS/model_resrit_etasexinteraction.jpg', height = 4, width=7)

```

Age x Sex Interaction
```{r}
p3 = ggplot(data = dat, aes(x = AGE, y = RIT_RES_ETA1, group_by(male2), color = male2)) +
  geom_smooth(aes(y = resrit_yhat_eta), method = 'lm') + 
  ylab('Higher-order Factor Scores') +
  xlab('Age (months)') +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_y_continuous(limits = c(-2,2)) +
  scale_color_discrete(name = "Sex", labels = c("Female", "Male")) +
  theme_bw() + 
  theme(text = element_text(size=16)) + 
  scale_x_continuous(breaks = seq(9, 36,3))
p3
```



